---
title: "Transform Data"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup}
library(dplyr)
library(babynames)
```

## Your Turn

Use summarise() to compute three statistics about the data:

1. The first (minimum) year in the dataset  
2. The last (maximum) year in the dataset  
3. The total number of children represented in the data

```{r}

```

## Your Turn

Extract the rows where `name == "Khaleesi"`. Then use `summarise()` and a summary functions to find:

1. The total number of children named Khaleesi
2. The first year Khaleesi appeared in the data

```{r}

```

## Your Turn

Use `group_by()`, `summarise()`, and `arrange()` to display the ten most popular names. Compute popularity as the total number of children of a single gender given a name.

```{r}

```

## Your Turn

Use grouping to calculate and then plot the number of children born each year over time.

```{r}

```

## Your Turn

Use `min_rank()` and `mutate()` to rank each row in `babynames` from largest `n` to lowest `n`.

```{r}

```

## Your Turn

Compute each name's rank _within its year and sex_. 
Then compute the median rank _for each combination of name and sex_, and arrange the results from highest median rank to lowest.



## Joining data

```{r}
library(nycflights13)
```

## Your turn
Read in the toy datasets band and instrument


## Types of joins

```{r}
band %>% left_join(instrument, by = "name")
band %>% right_join(instrument, by = "name")
band %>% full_join(instrument, by = "name")
band %>% inner_join(instrument, by = "name")
```

## Your turn
Which airlines had the largest arrival delays? Work in groups to complete the code below. 

```{r}
flights %>%
  drop_na(arr_delay) %>%
     #something!     %>%
  group_by(   #something!  ) %>%
     #something!      %>%
  arrange(  #something! )       
```

## Your turn
Read in the toy dataset instrument2

## What if the names don't match?

```{r}
band %>% left_join(instrument2, by = c("name" = "artist"))
```

```{r}
airports %>% left_join(flights, by = c("faa" = "dest"))
```


# Take aways

* Extract variables with `select()`  
* Extract cases with `filter()`  
* Arrange cases, with `arrange()`  

* Make tables of summaries with `summarise()`  
* Make new variables, with `mutate()`  
* Do groupwise operations with `group_by()`

* Connect operations with `%>%`  

* Use `left_join()`, `right_join()`, `full_join()`, or `inner_join()` to join datasets
* Use `semi_join()` or `anti_join()` to filter datasets against each other



