---
title: "R Structures"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Arithmetic and algebra

Let's start with the same arithmetic we were just doing in the Console. Notice that the math is surrounded by some special characters above and below, and has a grey background. This is RStudio's way of showing you the difference between an area to write text (here!) and an area where code is stored (called a code chunk).

```{r}
5 + 5
4 - 1
1 * 2
4 ^ 2
```

Try executing this chunk by clicking the green *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*. These two actions do the same thing-- they run all the lines of code in a chunk, and show you all the output.

You can also run just one line at a time by putting your cursor on a line and pressing *Cmd+Enter*. Then, you'll just see the output for the one line. 

We can do the same with the algebra we just saw,

```{r}
a <- 1
b <- 2
a + b
A <- 3
a + b - A
```

## Functions 

```{r}
round(3.1415)
factorial(3)
sqrt(9)

factorial(round(2.0015) + 1)
```


```{r}
the_answer <- 42
more_pi <- round(3.1415) + 1
factorial(more_pi)
```

What happened in your Environment when you ran that code? 

## "knitted" HTML version

This functionality is supported by an R package called `knitr`, so we often talk about "knitted" versions of RMarkdown and R Notebook documents. 

When you save the notebook, an HTML file containing the code and output will be saved alongside it. This makes a nice, polished report of your work to share.

Click the *Preview* button at the top of this document or press *Cmd+Shift+K* (*Ctrl+Shift+K* on Windows) to preview the HTML file. Try clicking *Preview* now.

## More on Notebooks

We'll be covering many more details about notebooks as we progress, but ask a TA if you have pressing questions right now. 

