---
title: "Syntax"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r setup}
library(ggplot2)
library(fivethirtyeight)
```

## Syntax in ggplot2
There are many ways to "say" the same thing in ggplot2

```{r}
ggplot(data = bechdel) +
  geom_point(mapping = aes(x = budget, y = domgross)) +
  geom_smooth(mapping = aes(x= budget, y = domgross))
```

```{r}
ggplot(data = bechdel, aes(x= budget, y = domgross)) +
  geom_point() +
  geom_smooth()
```

```{r}
ggplot(bechdel) + 
  geom_point(aes(x = budget, y = domgross))

ggplot(bechdel, aes(x = budget, y = domgross)) +
  geom_point()

ggplot(bechdel, aes(x= budget)) + 
  geom_point(aes(y = domgross))
```

Override the global mappings

```{r}
ggplot(data = bechdel, aes(x= budget, y = domgross)) +
  geom_point(aes(color=clean_test)) +
  geom_smooth()
```

## Subsetting

## The dplyr way
To use the dplyr package, we need to load it.

```{r}
library(dplyr)
```



## Beatles data
Let's make a toy dataset to play with. 

```{r}
beatles <- data.frame(
  name = c("John", "Paul", "George", "Ringo"),
  birth = c(1940, 1942, 1943, 1940), 
  instrument = c("guitar", "bass", "guitar", "drums")
)
```

## Select
Select is a way to extract columns from dataframes

```{r}
select(beatles, name, birth)
```

## Your turn
Select just the instrument column



## Base R
```{r}
beatles[2,3]
beatles[ ,"birth"]
beatles[ ,c("name","birth")]
beatles$birth
```

## Your turn
Select just the instrument column using brackets

```{r}

```

Select just the instrument column using a dollar sign.

```{r}

```

## Filter

```{r}
filter(beatles, name == "George")
```

## Your turn


## More than one condition
```{r}
filter(beatles, birth==1940, instrument == "guitar")
filter(beatles, birth==1940 & instrument == "guitar")
```

## Your turn
Modify  the code below to filter out the rows for which birth is 1943 or instrument is drums

```{r}
filter(beatles, birth==1940 & instrument == "guitar")
```


## Base R

```{r}
beatles[c(FALSE,TRUE,TRUE,FALSE), ]
beatles[beatles$birth == 1940, ]
```

## Your turn

## Baby names data
Names with at least 5 uses in a particular year, from the Social Security Administration. 

```{r}
library(babynames)
```

## Your turn
Do you remember how to view a dataset?

## Saving results

```{r}
filter(babynames, name %in% c("Acura", "Lexus", "Yugo"))
carnames <- filter(babynames, name %in% c("Acura", "Lexus", "Yugo"))
```

```{r}
babynames <- filter(babynames, name %in% c("Acura", "Lexus", "Yugo")) # dangerous!
rm(babynames)
str(babynames) #phew!
```

## Your turn
Try to:
- Filter out the babynames data for your name. (You may want to also filter for your gender)
- Assign that data its own name 
- Use your subsetted data to create a ggplot graphic of the popularity of your name over time

## Steps

```{r}
boys_2015 <- filter(babynames, year == 2015, sex == "M")
boys_2015 <- select(boys_2015, name, n)
boys_2015 <- arrange(boys_2015, desc(n))
boys_2015
```

Nested nonsense

```{r}
arrange(select(filter(babynames, year == 2015, 
  sex == "M"), name, n), desc(n))
```

## Piping
The `%>%` operator is the pipe. We can reorganized our code like so:

```{r}
babynames %>%
  filter(year == 2015, sex == "M") %>%
  select(name, n) %>%
  arrange(desc(n))
```

[Tip: type Cmd Shift M to insert a pipe]

## Your turn
Use %>% to write a sequence of functions that: 
- Filter babynames to just the girls that were born in 2015
- Select the `name` and `n` columns
- Arrange the results so that the most popular names are near the top.

```{r}

```

## Diamonds data

```{r}

```

