---
title: "Models - solutions"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, message=FALSE}
library(dplyr)
library(modelr)
library(ggplot2)
library(skimr)
library(broom)
wages <- heights %>% filter(income > 0)
```

## Wages data
```{r}
wages %>% skim()
wages %>% 
  ggplot(aes(log(income))) + geom_histogram(binwidth = 0.25)
```

## Linear models

```{r}
mod_e <- lm(log(income) ~ education, data = wages)
```

## Your Turn

Fit the model on the slide and then examine the output. What does it look like?

```{r}
mod_e
summary(mod_e)
names(mod_e)
mod_e$coefficients
```

## Plotting models with base R
```{r}
plot(mod_e)
plot(mod_e, which=c(1,2))
```

## With ggplot2

```{r}
ggplot(mod_e, aes(x=.fitted, y=.resid)) + geom_point() + geom_smooth(se = FALSE)

ggplot(mod_e, aes(sample = .resid)) + geom_qq()
```


## Broom
```{r}
mod_e %>% tidy()
mod_e %>% glance()
mod_e %>% augment()
mod_e %>% augment(data = wages)
```

## Your turn
Model log(income) against height. Then use broom and dplyr functions to extract:
The coefficient estimates and their related statistics
The adj.r.squared and p.value for the overall model

```{r}
mod_h <- lm(log(income) ~ height, data=wages)
mod_h %>%
  tidy()
mod_h %>%
  glance() %>%
  select(p.value, r.squared)
```

## Multivariate regression

## Your Turn

Model `log(income)` against `education` _and_ `height`. Do the coefficients change?

```{r}
mod_eh <- lm(log(income) ~ education + height, data = wages)

mod_eh %>% 
  tidy()
```


## Your Turn

Model `log(income)` against `education` and `height` and `sex`. Can you interpret the coefficients?

```{r}
mod_ehs <- lm(log(income) ~ education + height + sex, data = wages)

mod_ehs %>% 
  tidy()
```


## Logistic regression

```{r}
library(fivethirtyeight)
data(bechdel)
bechdel <- bechdel %>%
  mutate(pass = if_else(binary == "PASS", 0, 1))
```

```{r}
mod_pass <- glm(pass~budget, data=bechdel, family=binomial)
summary(mod_pass)
```

## Your turn
Model pass against budget, year and domgross_2013. `tidy()` the model output.

```{r}
mod_pass2 <- glm(pass~budget + year + domgross_2013, data=bechdel, family=binomial)

mod_pass2 %>%
  tidy()
```

***

# Take Aways

* Use `glance()`, `tidy()`, and `augment()` from the **broom** package to return model values in a data frame.

* Use `add_predictions()` or `gather_predictions()` or `spread_predictions()` from the **modelr** package to visualize predictions.

* Use `add_residuals()` or `gather_residuals()` or `spread_residuals()` from the **modelr** package to visualize residuals.


