---
title: "Visualization - solutions"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Setup

The first chunk in an R Notebook is usually titled "setup," and by convention includes the R packages you want to load. Remember, in order to use an R package you have to run some `library()` code every session. Execute these lines of code to load the packages. 

```{r setup}
library(ggplot2)
library(fivethirtyeight)
```

## Bechdel test data

We're going to start by playing with data collected by the website FiveThirtyEight on movies and [the Bechdel test](https://en.wikipedia.org/wiki/Bechdel_test). 

To begin, let's just preview our data. There are a couple ways to do that. One is just to type the name of the data and execute it like a piece of code. 

```{r}
bechdel
```

Notice that you can page through to see more of the dataset. 

Sometimes, people prefer to see their data in a more spreadsheet-like format, and RStudio provides a way to do that. Go to the Console and type `View(bechdel)` to see the data preview. 

(An aside-- `View` is a special function. Since it makes something happen in the RStudio interface, it doesn't work properly in R Notebooks. Most R functions have names that start with lowercase letters, so the uppercase "V" is there to remind you of its special status.)



## Visualizing data with ggplot2

## Your Turn 1

Run the code on the slide to make a graph. Pay strict attention to spelling, capitalization, and parentheses!

```{r}
ggplot(data = bechdel) +
  geom_point(mapping = aes(x = budget, y = domgross))
```

## Your Turn 2

Add `color`, `size`, `alpha`, and `shape` aesthetics to your graph. Experiment.  

```{r}
ggplot(data = bechdel) +
  geom_point(mapping = aes(x = budget, y = domgross, color=clean_test))
```

## Your Turn 3

Replace this scatterplot with one that draws boxplots. Use the cheatsheet. Try your best guess.

```{r}
ggplot(bechdel) + geom_boxplot(aes(clean_test, budget))
```

## Your Turn 4

Make a histogram of the `budget` variable from `bechdel`.

```{r}
ggplot(bechdel) + geom_histogram(aes(x=budget))
```

## Your Turn 5
Try to find a better binwidth for `budget.'

```{r}
ggplot(data = bechdel) +
  geom_histogram(mapping = aes(x = budget), binwidth=10000000)
```

## Your Turn 6

Make a density plot of `budget` colored by `clean_test`.

```{r}
ggplot(data = bechdel) +
  geom_density(mapping = aes(x = budget, color=clean_test))
```


## Your Turn 7

Predict what this code will do. Then run it.

```{r}
ggplot(bechdel) + 
  geom_point(aes(budget, domgross)) +
  geom_smooth(aes(budget, domgross))
```

## Your Turn 8

What does `getwd()` return?

```{r}
getwd()
```

## Your Turn 9

Save the last plot and then locate it in the files pane.

```{r}
ggsave()
```

***

# Take aways

You can use this code template to make thousands of graphs with **ggplot2**.

```{r eval = FALSE}
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```