---
title: "Syntax - Solutions"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Syntax in ggplot2
There are many ways to "say" the same thing in ggplot2

```{r}
ggplot(data = bechdel) +
  geom_point(mapping = aes(x = budget, y = domgross)) +
  geom_smooth(mapping = aes(x= budget, y = domgross))
```

```{r}
ggplot(data = bechdel, aes(x= budget, y = domgross)) +
  geom_point() +
  geom_smooth()
```

```{r}
ggplot(bechdel) + 
  geom_point(aes(x = budget, y = domgross))
ggplot(bechdel, aes(x = budget, y = domgross)) +
  geom_point()
ggplot(bechdel, aes(x= budget)) + 
  geom_point(aes(y = domgross))
ggplot() + 
  geom_point(bechdel, aes(x = budget, y = domgross))
```

Override the global mappings

```{r}
ggplot(data = bechdel, aes(x= budget, y = domgross)) +
  geom_point(aes(color=clean_test)) +
  geom_smooth()
```

## Subsetting

## The dplyr way
To use the dplyr package, we need to load it.

```{r setup}
library(dplyr)
```



## Beatles data
Let's make a toy dataset to play with. 

```{r}
beatles <- data.frame(
  name = c("John", "Paul", "George", "Ringo"),
  birth = c(1940, 1942, 1943, 1940), 
  instrument = c("guitar", "bass", "guitar", "drums")
)
```

## Select
Select is a way to extract columns from dataframes

```{r}
select(beatles, name, birth)
```

## Your Turn
Select just the instrument column

```{r}
select(beatles, instrument)
```

## Base R
```{r}
beatles[2,3]
beatles[ ,"birth"]
beatles[ ,c("name","birth")]
beatles$birth
```

## Filter

```{r}
filter(beatles, name == "George")
```

## Quiz

```{r}
x <- c(1, 2, 3, 4, 5)
x > 3
x >= 3
x < 3
x <= 3
x == 3
x != 3
x = 3
```

## Your turn

```{r}
filter(beatles, birth==1940)
```




## More than one condition
```{r}
filter(beatles, birth==1940, instrument == "guitar")
filter(beatles, birth==1940 & instrument == "guitar")
```

## Base R

```{r}
beatles[c(FALSE,TRUE,TRUE,FALSE), ]
beatles[beatles$birth == 1940, ]
```

## Baby names data
Names with at least 5 uses in a particular year, from the Social Security Administration. 

```{r}
library(babynames)
```

## Your turn
Do you remember how to view a dataset?

```{r}
filter(babynames, prop >= 0.08)
filter(babynames, name == "Sea")
filter(babynames, is.na(n))
filter(babynames, name == "Sea", sex == "F")
filter(babynames, n == 5 | n == 6, year == 1880)
filter(babynames, name %in% c("Acura", "Lexus", "Yugo"))
```

## Saving results

```{r}
filter(babynames, name %in% c("Acura", "Lexus", "Yugo"))
carnames <- filter(babynames, name %in% c("Acura", "Lexus", "Yugo"))
```

```{r}
babynames <- filter(babynames, name %in% c("Acura", "Lexus", "Yugo")) # dangerous!
rm(babynames)
str(babynames) #phew!
```

## Your turn 
```{r}
amelia <- babynames %>%
    filter(name=="Amelia" & sex == "F")

ggplot(amelia) + geom_point(aes(x=year, y=n))

ggplot(amelia) + geom_line(aes(x=year, y=n))
```
